### 2021-01-16 paper/localpsf.tex
more edits

\section{Constructing the approximation}
\label{sec:method}
This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\widetilde{\mathcal{A}} \approx \mathcal{A}$, given in \eqref{eq:product_convolution}. We also describe how we evaluate kernel entries of the approximation, $\widetilde{A}(y,x)$, and how we convert $\widetilde{\mathcal{A}}$ to H-matrix format.
\begin{itemize}
\item We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}). To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as distinct functions.
\item We use a greedy algorithm to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}). 
\item We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\psi_x$ (Section \ref{eq:mean_and_covariance_estimation}). 
\item For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
\end{itemize}
The complete algorithm for constructing $\widetilde{\mathcal{A}}$ is shown in Algorithm \ref{alg:construct_Atilde}. The method for evaluating kernel entries of $\widetilde{A}$ is shown in Section \ref{sec:eval_matrix_entries}. The method for converting $\widetilde{\mathcal{A}}$ to H-matrix format is shown in Section \ref{sec:H_matrix_conversion}. In the rest of the section we will need the following definitions.

\begin{defn}
	We define the spatially varying volume, $\alpha$, the scaled impulse response, $\rho_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
	\begin{align*}
	\alpha:\Omega \rightarrow \mathbb{R}, &\qquad \alpha(x) := \int_{\Omega} \phi_x(z) dz, \\
	\rho_x:\Omega \rightarrow \mathbb{R}, &\qquad \rho_x := \phi_x \big/ \alpha(x), \\
	\mu:\Omega \rightarrow \mathbb{R}^d, &\qquad \mu(x) := \int_\Omega z \rho_x(z) dz, \\
	\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\Sigma(x) := \int_\Omega (z - \mu(x))(z - \mu(x))^T \rho_x(z) dz.
	\end{align*}
\end{defn}

@@

Since $\phi_x$ is nonnegative, it is a scaled probability distribution. We compute the scaling factor, denoted $\alpha(x)$, for all $x$ simultaneously by applying $\mathcal{A}$ to a constant function. We then compute the spatially varying mean, $\mu(x)$, and covariance, $\Sigma(x)$ of the normalized version of $\phi_x$, for all $x$ simultaneously, by applying $\mathcal{A}^T$ to a small number of linear and quadratic functions. We make the approximation that the support of $\phi_x$ is contained within the ellipsoid
\begin{equation}
\label{eq:support_ellipsoid}
E_x := \{x' \in \Omega: (x' - \mu(x))^T \Sigma(x)^{-1} (x' - \mu(x)) \le \tau^2\},
\end{equation}
where $\tau$ is a small number of standard deviations. The bigger $\tau$ is, the smaller the overlap between impulse responses within a batch will be; the smaller $\tau$ is, the more impulse responses we can fit within each batch. In practice we found that $\tau=3$ provides a good balance between these competing goals. 
%The ellipsoid $E_{x}$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian approximation of the normalized version of $\langle \mathcal{A}, \delta_{x}\rangle^{*}$ which has the same mean and covariance. 
The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$. Note that this Gaussian distribution has the same mean and covariance as  which has the same mean and covariance as the normalized version of $\phi_x$ (see Figure FIG). 
The fraction of the mass of $\phi_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though we find this bound is conservative and far less mass resides in this region in practice.

>>>>

\section{Constructing the approximation}
\label{sec:method}
This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\widetilde{\mathcal{A}} \approx \mathcal{A}$, given in \eqref{eq:product_convolution}. We also describe how we evaluate kernel entries of the approximation, $\widetilde{A}(y,x)$, and how we convert $\widetilde{\mathcal{A}}$ to H-matrix format.
\begin{itemize}
\item We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}). To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as distinct functions.
\item We use a greedy algorithm to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}). 
\item We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\psi_x$ (Section \ref{eq:mean_and_covariance_estimation}). 
\item For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
\end{itemize}
The complete algorithm for constructing $\widetilde{\mathcal{A}}$ is shown in Algorithm \ref{alg:construct_Atilde}. The method for evaluating kernel entries of $\widetilde{A}$ is shown in Section \ref{sec:eval_matrix_entries}. The method for converting $\widetilde{\mathcal{A}}$ to H-matrix format is shown in Section \ref{sec:H_matrix_conversion}.

@@

Since $\phi_x$ is nonnegative, it is a scaled probability distribution. We compute the scaling factor, denoted $\alpha(x)$, for all $x$ simultaneously by applying $\mathcal{A}^T$ to a constant function. We compute the spatially varying mean, $\mu(x)$, and covariance, $\Sigma(x)$ of the normalized version of $\phi_x$, for all $x$ simultaneously, by applying $\mathcal{A}^T$ to a small number of linear and quadratic functions. Definitions of $\alpha$, $\mu$, and $\Sigma$ are given in Definition \ref{defn:alpha_rho_mu_sigma}. The procedure for computing these quantities is shown in Algorithm \ref{alg:varhpi_mean_cov}, and the theoretical basis for this procedure is shown in Theorem \ref{thm:vol_mean_cov}.

We make the approximation that the support of $\phi_x$ is contained within the ellipsoid
\begin{equation}
\label{eq:support_ellipsoid}
E_x := \{x' \in \Omega: (x' - \mu(x))^T \Sigma(x)^{-1} (x' - \mu(x)) \le \tau^2\},
\end{equation}
where $\tau$ is a small number of standard deviations. 
%The ellipsoid $E_{x}$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian approximation of the normalized version of $\langle \mathcal{A}, \delta_{x}\rangle^{*}$ which has the same mean and covariance. 
The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$, i.e., the Gaussian distribution which has the same mean and covariance as the normalized version of $\phi_x$ (see Figure FIG). The bigger $\tau$ is, the smaller the overlap between impulse responses within a batch will be, while the smaller $\tau$ is, the more impulse responses we can fit within each batch. In practice we found that $\tau=3$ provides a good balance between these competing goals. 
The fraction of the mass of $\phi_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though we find this bound is conservative and far less mass resides in this region in practice.

\begin{defn}
	\label{defn:alpha_rho_mu_sigma}
	We define the spatially varying volume, $\alpha$, the scaled impulse response, $\rho_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
	\begin{align*}
	\alpha:\Omega \rightarrow \mathbb{R}, &\qquad \alpha(x) := \int_{\Omega} \phi_x(z) dz, \\
	\rho_x:\Omega \rightarrow \mathbb{R}, &\qquad \rho_x := \phi_x \big/ \alpha(x), \\
	\mu:\Omega \rightarrow \mathbb{R}^d, &\qquad \mu(x) := \int_\Omega z \rho_x(z) dz, \\
	\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\Sigma(x) := \int_\Omega (z - \mu(x))(z - \mu(x))^T \rho_x(z) dz.
	\end{align*}
\end{defn}
